comgen-outline

### 2017-06-10 paper/comgen-paper.Rmd
habituals structured prior model, data, analysis and figure

```{r habituals-prior}
annualRates = list("5 years" = 1/5, "2 years" = 1/2,
                   "year" = 1, "6 months" = 2, "2 months" = 6,
                   "month" = 12, "2 weeks" = 26 ,"week" = 52)

d.hab.priors <- read.csv(paste(project.path,
                           "data/habituals/priors/friends-and-family-2-trials.csv",
                           sep= ""))

d.hab.priors.tidy <- d.hab.priors %>%
  rowwise() %>%
  mutate(annualRate = annualRates[[as.character(interval)]]*n_times,
         never = ifelse(annualRate == 0, 1, 0),
         action = as.character(action),
         action = ifelse(action == "sell things on eBay", "sells things on eBay", action))

d.hab.priors.distinctiveness <- d.hab.priors.tidy %>%
  group_by(action) %>%
  multi_boot_standard(column = 'never') %>%
  ungroup() %>%
  rename(distinctiveness = mean, 
         distinct_low = ci_lower, distinct_high = ci_upper)
#d.hab.priors
#d.hab.priors.distinctiveness
# 
# 
# df.out <- df.tidy %>%
#   mutate(annualRate = ifelse(annualRate == 0, 0.15, annualRate),
#          lograte = log(annualRate),
#          binned_lograte = round(lograte*2)/2) %>%
#   group_by(action, binned_lograte) %>%
#   summarize(counts = n())
# 
# 
# # make sure 0s are represented
# df.out2 <- left_join(
#   expand.grid(action = levels(df.out$action), binned_lograte = levels(factor(df.out$binned_lograte))),
#   df.out %>% mutate(binned_lograte = factor(binned_lograte))) 
# 
# df.out2[is.na(df.out2)] <- 0
# df.out2 <- df.out2 %>%
#   mutate(binned_lograte = as.numeric(binned_lograte))
#   
# ggplot(df.out2, aes( x = binned_lograte, y = counts))+
#   #geom_histogram()+
#   geom_bar(stat = 'identity', position = position_dodge())+
#   #geom_density()+
#   facet_wrap(~action, scales = 'free')+
#   xlim(-3.5, 8.5)
# d.hab.priors.tidy
```

>>>>

```{r habituals-prior-data}
annualRates = list("5 years" = 1/5, "2 years" = 1/2,
                   "year" = 1, "6 months" = 2, "2 months" = 6,
                   "month" = 12, "2 weeks" = 26 ,"week" = 52)

d.hab.priors <- read.csv(paste(project.path,
                           "data/habituals/priors/friends-and-family-2-trials.csv",
                           sep= ""))
d.hab.priors <- read.csv("/Users/mht/Documents/research/habituals/data/priors/priors-2.csv")

d.hab.priors <- d.hab.priors %>%
  mutate(comparisonTime_men = as.character(comparisonTime_men),
         comparisonTime_women = as.character(comparisonTime_women)) %>%
  rowwise() %>%
  mutate(
    mixture_male = nPersons_men / comparisonNum_men,
    mixture_female = nPersons_women / comparisonNum_women,
    dayRate_male = ifelse(comparisonTime_men == "week", nInstances_men / 7,
                   ifelse(comparisonTime_men == "month", nInstances_men / 30,
                   ifelse(comparisonTime_men == "year", nInstances_men / 365,
                   ifelse(comparisonTime_men == "5 years", 
                          nInstances_men / (5*365), -99)))),
    dayRate_female = ifelse(comparisonTime_women == "week", nInstances_women / 7,
                   ifelse(comparisonTime_women == "month", nInstances_women / 30,
                   ifelse(comparisonTime_women == "year", nInstances_women / 365,
                   ifelse(comparisonTime_women == "5 years", 
                          nInstances_women / (5*365), -99)))),
    annualRate_male = dayRate_male * 365,
    annualRate_female = dayRate_female * 365
  )

d.hab.priors.filtered <- filter(d.hab.priors, 
       !(
         (mixture_male == 0) ||   (mixture_female == 0) ||   
           (dayRate_male == 0) ||   (dayRate_female == 0)  )
       )

```{r habituals-prior-forwardsample}
# this is basically bootstrapping and combining using the structured model

d.hab.priors.samples <- data.frame()
for (i in 1:10){

  d.hab.priors.samples <- bind_rows(
    d.hab.priors.samples,
    d.hab.priors.filtered %>% 
    select(workerid, item, starts_with("annualRate"), 
           starts_with("mixture")) %>%
    rowwise() %>%
    mutate(
      stable_male = rbinom(n = 1, size = 1, prob = mixture_male),
       freq_male = ifelse(stable_male == 1, annualRate_male, 0.007),
       stable_female = rbinom(n = 1, size = 1, prob = mixture_female),
       freq_female = ifelse(stable_female == 1, annualRate_female, 0.007)
       ) %>%
    select(item, freq_male, freq_female) %>%
    gather(gender, val, freq_male, freq_female) %>%
  mutate(gender = gsub("freq_", "", gender))
  )
  print(i)
}

```{r habituals-prior-model}
hab.prior.bda.model <- '
var betaShape = function(p){
  return {a: p.g * p.d, b: (1-p.g) * p.d}
};

var model = function(){
	var mixtureParams = {
		male: {
      g: uniformDrift({a: 0, b: 1, width: 0.2}),
      d: uniformDrift({a: 0, b: 100, width: 5})
    },
		female: {
      g: uniformDrift({a: 0, b: 1, width: 0.2}),
      d: uniformDrift({a: 0, b: 100, width: 5})
    }
	};

	var mixtureShapes = {
		male: betaShape(mixtureParams.male),
		female: betaShape(mixtureParams.female)
	};

	mapData({data: data.mixture}, function(d){
    Beta(mixtureShapes[d.gender]).score(d.val) == -Infinity ? display(JSON.stringify(d)) : null
     // display(Beta(mixtureShapes[d.gender]).score(d.val))
		observe(Beta(mixtureShapes[d.gender]), d.val)
	})

	var stableFrequency = {
		male: {
			mu: uniformDrift({a: -5, b:10, width: 2}),
			sigma: uniformDrift({a:0, b:5, width: 1})
		},
		female: {
			mu: uniformDrift({a:-5, b:10, width: 2}),
			sigma: uniformDrift({a:0, b:5, width: 1})
		}
	}

	mapData({data: data.frequency}, function(d){
    Gaussian(stableFrequency[d.gender]).score(d.logval) == -Infinity ? 
        display(JSON.stringify(d)) : null
    // display(Gaussian(stableFrequency[d.gender]).score(d.logval))
		observe(Gaussian(stableFrequency[d.gender]), d.logval)
	})

	var existenceProb = {
		male: beta(mixtureShapes.male),
		female: beta(mixtureShapes.female)
	};

	var freqWhenPresent = {
		male: gaussian(stableFrequency.male),
		female: gaussian(stableFrequency.female)
	}

  var marginalFreq = {
    male: flip(existenceProb.male) ? freqWhenPresent.male : -5,
    female: flip(existenceProb.female) ? freqWhenPresent.female : -5,
  }

  return {
    mix_mean_male : mixtureParams.male.g,
    mix_mean_female : mixtureParams.female.g,
    mix_samplesize_male : mixtureParams.male.d,
    mix_samplesize_female : mixtureParams.female.d,
    stableFreq_mean_male : stableFrequency.male.mu,
    stableFreq_mean_female :stableFrequency.female.mu,
    stableFreq_sd_male : stableFrequency.male.sigma,
    stableFreq_sd_female :stableFrequency.female.sigma,
    marginalFreq_NA_male: marginalFreq.male,
    marginalFreq_NA_female: marginalFreq.female
  }
}
'
```

```{r habituals-prior-model-run}
items <- levels(d.hab.priors.filtered$item)
n_samples <- 500
rs.habituals.bda.prior <- data.frame()

for (it in items){

 df.prior.toPass <- list(
   frequency = d.hab.priors.filtered %>%
      filter(item == it) %>%
      select(item, starts_with("annualRate")) %>%
      gather(gender, val, -item) %>%
      mutate(gender = gsub("annualRate_", "", gender),
             logval = log(val)),
   mixture = d.hab.priors.filtered %>% 
      filter(item == it) %>%
      select(item, starts_with("mixture")) %>%
      gather(gender, val, -item) %>%
      mutate(gender = gsub("mixture_", "", gender)) %>%
      rowwise() %>%
      mutate(val = ifelse(val == 1, 0.9999, 
                          ifelse(val == 0, 0.000001, val)))
  )

  rs <- webppl(program_code = hab.prior.bda.model,
       model_var = "model",
       inference_opts = list(method = "MCMC", samples = n_samples,
                             burn = n_samples / 2),
       data = df.prior.toPass,
       data_var = "data")

  rs.habituals.bda.prior <- bind_rows(
    rs.habituals.bda.prior, 
    rs %>% 
      mutate(item = it) %>% 
      separate(Parameter, into = c("Variable", "Parameter", "Gender"))
  )
}

rs.habituals.bda.prior %>% 
  mutate(param = paste(Variable, Parameter, sep = "_")) %>%
  select(-Variable, -Parameter, -Chain) %>%
  spread(param, value) 
```

```{r habituals-prior-model-forwardSample}
rs.habituals.bda.prior.spread <- rs.habituals.bda.prior %>% 
  mutate(param = paste(Variable, Parameter, sep = "_")) %>%
  select(-Variable, -Parameter, -Chain) %>%
  spread(param, value)

rs.habituals.bda.prior.samples <- data.frame()
for (i in 1:1){
  rs.habituals.bda.prior.samples <- bind_rows(
    rs.habituals.bda.prior.samples,
    rs.habituals.bda.prior.spread %>%
    rowwise() %>%
    mutate(
      a = mix_mean * mix_samplesize,
      b = (1 - mix_mean) * mix_samplesize,
      theta = rbeta(n = 1, shape1 = a, shape2 = b),
      stable = rbinom(n = 1, size = 1, prob = theta),
      logannualRate = ifelse(stable == 1, 
                             rnorm(n = 1, 
                                   mean = stableFreq_mean, 
                                   sd = stableFreq_sd), -5),
       annualRate =  exp(logannualRate)
    )
  )
}
```

```{r habituals-prior-fig, fig.width = 8.5, fig.height = 4, fig.cap="Reconstructed priors on frequency from the structured, prior elicitation task (Expt. 2a). Priors are reconstructed by bootstrapping and forward sampling using the structured model (data, blue) or using Bayesian data analysis on the structured model (model, red)."}

bind_rows(
  d.hab.priors.samples %>% 
    mutate(src = 'data', 
           val = log(val)),
  rs.habituals.bda.prior.samples %>%
    rename(gender = Gender, val = logannualRate) %>%
    select(gender, item, val) %>%
    mutate(src = 'model')
  ) %>%
ggplot(. , aes( x = val, color = src))+
  geom_density(size = 0.8, aes( y = ..scaled.. ))+
  facet_wrap(~item, nrow = 4)+
  scale_color_solarized()+
  xlab("Frequency")+
  ylab("Scaled prior density")+
  scale_x_continuous(limits = c(-5,10)) +
  scale_y_continuous(limits = c(0,1), breaks = c(0, 0.5, 1)) +
  theme(strip.text.y = element_text(angle = 0),
        legend.position = c(0.95, 0.075),
        axis.text.x = element_text(angle = 90))+
    scale_x_continuous(limits = c(-5, 8), 
                     breaks = c(-5, 0, 2.5, 4, 5.9),
                     labels = c("almost never", "annually", "monthly", "weekly", "daily"))
